---
title: "Reproducible Benchmarks and Packaged Example Outputs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducible Benchmarks and Packaged Example Outputs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ppmWass)
library(readr)

example_dir <- function(...) {
  inst_path <- system.file(..., package = "ppmWass")
  if (nzchar(inst_path)) {
    return(inst_path)
  }
  candidates <- c(
    file.path("..", "inst", ...),
    file.path("inst", ...)
  )
  hit <- candidates[file.exists(candidates)][1]
  if (!is.na(hit) && nzchar(hit)) {
    return(hit)
  }
  candidates[1]
}
```

This vignette shows how to inspect the representative benchmark outputs bundled
with the package and how to regenerate them for your own data.

The packaged artifacts are synthetic. They are intended as lightweight
templates for manuscript tables and supplementary files, not as scientific
evidence for method performance on external datasets.

The bundled examples are intentionally small. They are useful for checking file
formats, column names, and figure/table wiring, but manuscript claims should be
supported by larger real datasets run through the same scripts.

## Metric Cheat Sheet

- `replicate_top1`: fraction of replicate queries whose nearest neighbor shares the same InChIKey group.
- `replicate_mrr`: mean reciprocal rank of the first matching replicate across all replicate queries.
- `map`: mean average precision across ranked retrieval results.
- `auc_roc`: area under the ROC curve when InChIKey connectivity matches are treated as positives.
- `silhouette`: average silhouette width for cluster separation under the available labels.
- `fdr_at_07`: estimated false discovery rate among matches whose similarity exceeds `0.7`.
- `homolog_score`: internal consistency score for high-similarity, RI-separated pairs, not a curated homolog benchmark unless external labels are supplied.

## Bundled Benchmark Artifacts

```{r}
benchmark_dir <- example_dir("extdata", "benchmark_demo")
benchmark_summary <- read_csv(file.path(benchmark_dir, "benchmark_summary.csv"), show_col_types = FALSE)
benchmark_summary
```

The parameter-sweep example includes repeated subsampling (`n_replicates = 3`,
`sample_frac = 0.8`) so the output can be summarized into stability ranges
rather than treated as a single point estimate.

```{r}
sweep_summary <- read_csv(file.path(benchmark_dir, "distance_param_sweep.csv"), show_col_types = FALSE)
head(sweep_summary)
```

The bundled sweep was generated from a small grid like this:

```{r}
data.frame(
  tol_ppm = c(10, 10, 20, 20),
  w_frag = c(0.7, 0.5, 0.9, 0.6),
  w_loss = c(0.3, 0.5, 0.1, 0.4)
)
```

For manuscript work, expand this grid to cover the parameters that matter for
your data, then summarize the repeated runs by mean, SD, or quantiles.

Runtime summaries are also bundled so that the expected table structure is
available without rerunning the full benchmark workflow.

```{r}
runtime_summary <- read_csv(file.path(benchmark_dir, "distance_runtime_summary.csv"), show_col_types = FALSE)
runtime_summary
```

## Bundled Library Search Example

The package also ships a tiny MSP-based query/library example under
`inst/extdata/library_demo`.

```{r}
library_dir <- example_dir("extdata", "library_demo")
library_summary <- read_csv(file.path(library_dir, "library_benchmark_summary.csv"), show_col_types = FALSE)
library_summary
```

These files were generated from:

- `query_demo.msp`
- `library_demo.msp`
- `library_hits_cosine.csv`
- `library_hits_weighted_cosine.csv`
- `library_hits_composite.csv`

You can use them as templates for local NIST/MoNA-style MSP exports.

## Optional Distance Backends

The packaged example outputs focus on methods that do not require optional
dependencies beyond the core workflow. `entropy` requires the `msentropy`
package, and `wasserstein` / `ppm_wasserstein` require the `transport`
package, so those methods are not embedded in the small bundled CSV examples.

For paper figures or supplement tables, they should still be evaluated when the
optional dependencies are available:

```r
params <- eihrms_default_params()

run_benchmark(
  spectra_result = spectra,
  methods = c(
    "cosine", "weighted_cosine", "composite",
    "entropy", "wasserstein", "ppm_wasserstein"
  ),
  params = params
)
```

## Regenerating the Outputs

The repository includes two runnable scripts:

```r
# Full benchmark workflow
Rscript inst/scripts/run_benchmark_reproducible.R --output-dir=benchmark_output

# Query-vs-library retrieval workflow
Rscript inst/scripts/run_library_head_to_head.R \
  --query=query.msp \
  --library=library.msp \
  --output-dir=library_benchmark_output
```

For repository maintenance, the bundled synthetic example files are refreshed by:

```r
Rscript data-raw/generate_example_artifacts.R
```

## Interpretation Notes

- `evaluate_homolog_detection()` should be treated as an internal consistency metric when `compound_class` was inferred by the package itself.
- `create_homolog_ground_truth()` uses the first 14 characters of InChIKey, which represent the connectivity layer, not curated homolog-series labels.
- Full-dataset benchmark summaries are point estimates unless you also run repeated subsampling, bootstrap resampling, or another stability analysis.
